Hierarchical storage management in user space for neuroimaging applications

### 2025-10-15 paper/sea-neuro/paper-neuro.tex
Add more paper edits and updated figures

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

@@

    \subsection{Sea}

@@

    \vhs{It is important to note that Sea does not chunk files, and therefore, there is no risk of partitioned files across storage devices. Rather, it relies on configuration parameters
    \texttt{max\_fs} and \texttt{n\_threads} to determine whether there is sufficient space on a storage device to write the file to. \texttt{max\_fs} represents an upper bound estimate of the maximum
    filesize generated by the application in bytes. }

@@

    Sea can easily be executed directly using the available containers on the
    GitHub registry, or can be compiled via source using gcc-C++ and Make. Once
    compiled, Sea can be executed using the \texttt{sea\_launch.sh} bash script.

@@

    %% Add SPM here
    To preprocess the datasets with SPM, we reused the template described in
    \cite{haitas2021age}, with fieldmap correction removed. As the template was
    designed for a specific dataset, we updated the paths to the fMRI and
    anatomical images, as well as the dimensions, such as the number of volumes
    and slices and the TR and TA, for each dataset. Furthermore, we set the
    slice order to interleaved in all cases. SPM version 12 was used.

    For AFNI preprocessing, the pipeline was configured to perform slice timing,
    alignment to Talaraich space, registration, smoothing and brain masking.
    While AFNI does have some parallelized components that can be controlled via
    environment variables, we chose to let the pipeline use as much parallelism
    as required for its execution. AFNI version 21.1.02 was used.

@@

    To validate that the outputs produced were sufficiently correct, we
    performed visual quality control of the preprocessed images to ensure that
    they were correctly skull-stripped and registered to the template brain (MNI
    template). We repeated this step for all datasets and all pipelines
    processing a single image. In addition, we compared output number of files
    and output size to ensure that the pipeline was not producing different
    outputs depending on the file system used.

@@

Speedups were obtained by using Sea when Lustre
performance had been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}),
although performance was variable. The largest speedup across all conditions
was found to be 32$\times$ and occurred when SPM was processing a single
HCP image with 6 active busy writers.
On average, however, a single SPM pipeline
process preprocessing a single HCP image produced the greatest speedup
(12.6$\times$). The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
using a single process was the next fastest pipeline, with an average speedup of
(4.3$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
excellent speedups, which is likely due to a mix of prefetching the initial
input files and the I/O patterns of the application. While it was expected that the
AFNI pipeline would benefit the most from Sea due to its
minimal compute time and the amount of data it generates, it did not benefit
from Sea to the extent that the SPM pipeline did. As per Table~\ref{table:seaneuro-pipelines}, the
AFNI pipeline performs a very high number of glibc calls. While the overhead of an
individual call is likely very minimal, it is possible that the culmination of
all these calls results in a large enough bottleneck that affected both Sea and
Baseline executions to an extensive amount. Moreover, the number of glibc calls
aimed at interacting with data on the Lustre storage are very few. The AFNI pipeline still
obtained an average of 2$\times$ speedup across all experimental conditions.

>>>>

\definecolor{verylightgrey}{rgb}{0.92,0.92,0.92}
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

@@

    \subsection{The Sea data management library}

@@

    \vhs{It is important to note that Sea does not chunk files, and therefore, there is no risk of corruption due to partitioned files across storage devices. Rather, it relies on configuration parameters
    \texttt{max\_fs} and \texttt{n\_threads} to determine whether there is sufficient space on a storage device to write the file to. \texttt{max\_fs} represents an upper bound estimate of the maximum
    filesize generated by the application in bytes, whereas \texttt{n\_threads} is the maximum number of parallel threads that could be writing a file of size \texttt{max\_fs} concurrently. }

    \vhs{While Sea can function with applications that memory-map (memmap) files, it does not enhance the performance in instances where the file is memmapped. Memmap is an operation by which a small fragment of a large file is loaded (mapped) into memory. As the path of a memmaped file remains the same and data is updated in place, Sea will not be able to enhance performance of the I/O operations. }

@@

    \vhs{Sea can run alongside any application that dynamically loads and calls the libc library for read/write operations. Such applications commonly include interpreted applications, such as Python and Matlab/Octave. Sea can additionally work seamlessly from within a containerized environment, as long as the necessary cache directories required by Sea are mounted to the container. We provide various Docker containers preinstalled with Sea within the Github Container Registry (\url{https://github.com/orgs/big-data-lab-team/packages?repo_name=Sea}). Alternatively, Sea can be compiled from source using Make.} Once
    compiled, Sea can be executed using the \texttt{sea\_launch.sh} bash script.

        \begin{figure}
      \centering
       \begin{lstlisting}[language=bash,frame=single,
    backgroundcolor=\color{verylightgrey}, caption={Example Sea deployment command}, basicstyle=\ttfamily, label=lst:launch]
        $ SEA_HOME=$PWD ./sea_launch.sh <myprogram>
      \end{lstlisting}
    \end{figure}

@@

    %% Add SPM here
    To preprocess the datasets with SPM, we reused the template described in
    \cite{haitas2021age}, with fieldmap correction removed. \vhs{As with the other applications, MNI registration was performed.} Since the template was
    designed for a specific dataset, we updated the paths to the fMRI and
    anatomical images, as well as the dimensions, such as the number of volumes
    and slices and the TR and TA, for each dataset. Furthermore, we set the
    slice order to interleaved in all cases. SPM version 12 was used.

    For AFNI preprocessing, the pipeline was configured to perform slice timing, \vhs{EPI alignment to anatomical, 
    linear alignment to MNI space,} registration, smoothing \vhs{(blur size of 4)} and brain masking.
    While AFNI does have some parallelized components that can be controlled via
    environment variables, we chose to let the pipeline use as much parallelism
    as required for its execution. AFNI version 21.1.02 was used.

@@

    To validate that the outputs produced were \vhs{unaltered by Sea, we fixed the application seeds and}
    performed visual quality control of the preprocessed images to ensure that
    they were correctly skull-stripped and registered to the template brain (MNI
    template). We repeated this step for all datasets and all pipelines \vhs{(with and without using Sea)}
    processing a single image. In addition, we compared output number of files\vhs{, file hashes (MD5),} 
    and output size to ensure that the pipeline was not producing different
    outputs depending on the file system used.

@@

Speedups were obtained by using Sea when Lustre
performance had been degraded by busy writers (Figure~\ref{fig:seaneuro:slashbin}),
although performance was variable. The largest speedup across all conditions
was found to be 32$\times$ and occurred when SPM was processing a single
HCP image with 6 active busy writers.
On average, however, a single SPM pipeline
process preprocessing a single HCP image produced the greatest speedup
(12.6$\times$). The AFNI pipeline preprocessing a single fMRI image of the PREVENT-AD dataset
using a single process had the \vhs{next greatest speedup}, with an average speedup of
(4.3$\times$). The FSL Feat pipeline had speedups as well, with a maximum average speedup of
1.3$\times$ when preprocessing a single PREVENT-AD image. The SPM pipeline consistently had
excellent speedups, which is likely due to a mix of prefetching the initial
input files and the I/O patterns of the application. While it was expected that the
AFNI pipeline would benefit the most from Sea due to its
minimal compute time and the amount of data it generates, it did not benefit
from Sea to the extent that the SPM pipeline did. As per Table~\ref{table:seaneuro-pipelines}, the
AFNI pipeline performs a very high number of glibc calls. While the overhead of an
individual call is likely very minimal, it is possible that the culmination of
all these calls results in a large enough bottleneck that affected both Sea and
Baseline executions to an extensive amount. Moreover, the number of glibc calls
aimed at interacting with data on the Lustre storage are very few. The AFNI pipeline still
obtained an average of 2$\times$ speedup across all experimental conditions.
